---
title: "p8105_hw3_ma4197"
author: "Mayuri Albal"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library("scales")
```

##Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```


##Problem 2

```{r}
accel_data = read_csv(file= "./data/accel_data.csv") %>%
janitor::clean_names() %>%
  mutate(
    wkd_wkn= ifelse(day %in% c('Sunday', 'Saturday'), 'weekend', 'weekday')) %>% 
  relocate(week,day_id,day,wkd_wkn) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "active_num",
    names_prefix = "activity_",
    names_transform=list(active_num= as.numeric),
    values_to = "active_counts")
   
view(accel_data)
skimr::skim(accel_data)
```
In this cleaned dataset of 'accel_data' there are 6 variables: "week", "day_id", "day",
"wkd_wkn", "active_num", "active_count". "Active_num" and "active_count" were generated
from the previously displayed data for each activity that would occur (1-1440) and the 
corresponding count value the accelerometer measured. There are 50,400 observations within
this new data set.


###Aggregated Data
```{r}
accel_data %>% 
  group_by(day_id) %>% 
  mutate(
    tot_active= sum(active_counts)) %>% 
  summarize(day_id, tot_active) %>% 
  distinct %>% 
  print (n=)
```

####Explanation
Based on the printed table above, there are 35 days included in the data set. When observing
the total activity counts ("tot_active") across the 35 days, there does not appear to be
an definable trend in the total activity of the individual. However, there are dips in total
activity count on days 24 and 31.

###Graphing

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

###Plot
```{r}
ggplot(accel_data, aes(x= active_num, y= active_counts, color=day)) +
  geom_line(aes(group= day, color=day))+
  theme_classic()+ 
  theme(legend.position= "bottom")+
  ggtitle("24-Hour Activity Time Count")+
  theme(plot.title = element_text(hjust= 0.5))+
    labs(x = "Active Minutes",
        y= "Active Counts") +
  scale_x_continuous(breaks= c(0, 250, 500, 750, 1000, 1250, 1500))
   scale_colour_brewer(type = "seq", palette = "Spectral")
```


####Explanation
Based upon this graph, there appears to peaks in active counts being measured from the
accelerometer on Saturday and Sunday for this participant. Additionally, it appears that
during 1250 minutes throughout all of the days, there is a peak in active minutes around 7500 active counts. However for Saturday and Sunday, this peak is observed earlier in the daya during the 500 to 750 minutes as well as 1000 minutes.

##Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
view(ny_noaa)
skimr::skim(ny_noaa)
```
####Description
In the "ny_noaa" there are 7 variables and 2,595,176 observations. This data looks
at the 5 key variables that all New York weather stations assess, from January 1, 1981, to December 31st, 2010. Key variables include:'id', 'date', 'prcp', 'snow', 'snwd', 'tmax' and 'tmin'. In this data set, a significant portion of the data is missing from some of the variables. For prcp, about 145,838 observations are missing, with approximately 94.38% of the data having observations for the variable. For snow and snwd, approximately 85.31% and 77.19% of the data have observations for those variables. However for both tmax and tmin only 56.29%
of the data have observation  values for these variables, indicating that overall the missing data may bias the analysis due to the portion missing/labeled na.

###Part 1- Data Cleaning

```{r}
ny_noaa_3 =
ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, c("year", "month", "day"), sep = "-") %>%
  mutate(
    year= as.numeric(year),
    month= as.numeric(month),
    day=as.numeric(day),
   tmin= as.numeric(tmin),
   tmax= as.numeric(tmax),
   snow= as.numeric(snow),
    tmin= tmin/10,
    tmax = tmax/10,
    prcp = prcp/10
  )
skimr::skim(ny_noaa)
```

```{r}
ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, c("year", "month", "day"), sep = "-") %>%
  mutate(
    year= as.numeric(year),
    month= as.numeric(month),
    day = as.numeric(day),
    snow= as.numeric(snow),
    tmin= as.numeric(tmin),
    tmax= as.numeric(tmax),
    tmin= tmin/10,
    tmax = tmax/10,
    prcp = prcp/10
  )
```

###Snow Count
```{r}
count(ny_noaa_3, snow) 
```

####Explanation
From the original data set 'ny_noaa', which had 7 variables and 2,595176 observations, a new data set was formed called 'ny_noaa_3' with 9 variables and 2,595,176 observations. In this new data set, the variables are: 'id', 'year', 'month', 'day', 'prcp', 'snow', 'snwd',
'tmax', and 'tmin'. The variables 'year', 'month', and 'day' were formed from the variable 'date' in the original data set. The variables 'prcp', 'tmin', and 'tmax' were divided by 10
due to their orignal units being a tenth of a mm or Celcius, while the other variables were not, thus was done to ensure reasonable units. 

When observing the variable snowfall, the most commonly observed values were 0mm of snowfall and NA. This signifies that many days between January 1981 and December 2010 did not have snowfall in New York, which can be corroborated due to snowfall being only during winter months and under specific conditions, thus more likely to have more dates reporting NA or 0 mm of snowfall.

###Plot
```{r}
ggplot(ny_noaa_3, aes(x=year, y=tmax, color= month))+
  geom_point(alpha =0.5) +
facet_grid(.~month)
```


