---
title: "p8105_hw3_ma4197"
author: "Mayuri Albal"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
```

##Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```


##Problem 2

```{r}
accel_data = read_csv(file= "./data/accel_data.csv") %>%
janitor::clean_names() %>%
  mutate(
    wkd_wkn= ifelse(day %in% c('Sunday', 'Saturday'), 'weekend', 'weekday')) %>% 
  relocate(week,day_id, day,wkd_wkn) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "active_num",
    names_prefix = "activity_",
    names_transform=list(activity_=as.numeric),
    values_to = "active_counts") %>% 
    mutate( as.numeric("activity_"))
view(accel_data)
skimr::skim(accel_data)
```
In this cleaned dataset of 'accel_data' there are 6 variables: "week", "day_id", "day",
"wkd_wkn", "active_num", "active_count". "Active_num" and "active_count" were generated
from the previously displayed data for each activity that would occur (1-1440) and the 
corresponding count value the accelerometer measured. There are 50,400 observations within
this new data set.


###Aggregated Data
```{r}
accel_data %>% 
  group_by(day_id) %>% 
  mutate(
    tot_active= sum(active_counts)) %>% 
  summarize(day_id, tot_active) %>% 
  distinct
```

###2x2 Table